Do motif roles provide unique information about species risk of extinction?

### 2021-06-04 manuscript/motifs_vs_degree.tex
Updates from Overleaf

     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, eklof2006species}. Traditionally, there are two main approaches for studying secondary extinctions. First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}. Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models. However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.

@@

    %\subsection*{3 Global scale network properties are known to affect stability}
    The global structural properties of ecological networks have implications for their functioning \citep{petchey2002functional}, stability \citep{allesina2012stability} and robustness to disturbances \citep{dunne2002network, eklof2006species}. For example, more highly connected networks are more resistant to secondary extinctions \citep{dunne2002network, eklof2006species} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
    %\subsection*{4 meso-scale nw propoerties also affect stability}
    Structural properties on the local scale, i.e. properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}. However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance. On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.

@@

		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}. 
		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04). 
		For full details, see~\citet{Cirtwill2021_inprep}.
        We then filtered these simulated networks to remove any with unreasonable structures.
		Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where the shortest path between a consumer and the closest basal resource was \textgreater6 (as these high trophic levels are very uncommon in empirical food webs ~\citep{}).
		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.

>>>>

     %\subsection*{2 Bayesian networks are a fantastic way to model bottom-up effects}
    The study of how an initial decline or extinction leads to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{curtsdotter2011robustness, dunne2009cascading, Eklof2006}. Traditionally, there are two main approaches for studying secondary extinctions. First, there are topological models, which are solely based on food web structure \citep{dunne2009cascading}. Here, initial extinctions only affect other species in the network once a consumer has lost all of its prey and therefore must go extinct. The second approach uses dynamical models, which explicitly simulate population dynamics using a system of differential equations \citep{binzer2011susceptibility}. Dynamical models take changes in prey or predator densities into account when calculating the densities of other species in the network. These models therefore describe complex effects such as indirect interactions and are generally more realistic than topological models. However, dynamical models are parameter intensive, can only include a limited number of species, and simulations are time-consuming.

@@

    %\subsection*{3 Global scale network properties are known to affect stability}
    The global structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002d, Eklof2006}. For example, more highly connected networks are more resistant to secondary extinctions \citep{Dunne2002d, Eklof2006} and a modular organization where some groups of species are tightly interconnected with limited connections between groups can limit the propagation of disturbances \citep{}.
    %\subsection*{4 meso-scale nw propoerties also affect stability}
    Structural properties on the local scale, i.e. properties of single species, are also well-studied and there is consensus that species with high trophic levels, few prey (low in-degree), and low abundances are generally at higher risks of secondary extinction \citep{binzer2011susceptibility}. However, in-degree and trophic level are both relatively coarse descriptions of a species' role within its food-web~\citep{Cirtwill2018FoodWebs}. 
    For example, specialists may not necessarily have a high risk of extinction if it consumes a basal resource which is at low risk of being affected of disturbance. On the other hand, a consumer with low trophic level may still be at high risk of extinction if it consumes few preys or prey which possess high risk of being affected by disturbance.

@@

		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw,Delmas2017}. 
		To capture a range of plausible network architectures, we simulated networks with sizes ranging from 50 to 100 species (in steps of 10) and connectances ranging from 0.02 to 0.18 (in steps of 0.04). 
		For full details, see~\citet{Cirtwill2021_inprep}.
        We then filtered these simulated networks to remove any with unreasonable structures.
		Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where the shortest path between a consumer and the closest basal resource was \textgreater6 (as these high trophic levels are very uncommon in empirical food webs ~\citep{}).
		New networks were simulated to replace any removed networks, and this process was repeated until we obtained 100 suitable networks in each combination of size and connectance.
